---
title: "New York City AQI Data Exploration"
output: html_notebook
---

### Environment Setup

```{r warning=FALSE}
# Load Libraries
library(RSocrata)
library(aqsr)
library(usethis)
library(devtools)
library(data.table)
library(tidyverse)
library(DataExplorer)
library(filenamer)
library(validate)
library(request)
library(downloader)
```

## Dataset URLs

We have compiled a list of data sets we'd like to explore for this analysis. Each URL contains a query to filter down the records between Jan 1 2018 and Jan 1 2022, and in Manhattan borough or New York county, or New York City.

```{r}
source("./api/nyc_dataset_urls.R")
```

## Download Data From URLs

Download the URLs onto disk, to avoid storing in R memory. Taxi records are >100+ million, so downloading on disk would be faster. You can see status of download in print comments.

```{r warning=FALSE}
#source("./api/nyc_dataset_downloads.R")
dataset <- readRDS("./api/processed/nyc_dataset_list.csv")
```

# Exploratory Data Analysis

```{r}
# store data
dataset$loc

for (i in 1:nrow(dataset)){
  name <- dataset[i,]$feature_name
  readRDS(name, dataset[i,]$loc)
  
  print("Dataset saved successfully!")
}
```

```{r}
head(fhv1_19)
```

```{r}
head(data)
create_report(data)
```

```{r}
features <- profile_missing(data) %>% filter(pct_missing < 0.35)
features$feature
```

```{r}
streetConstructionPermits <- data[features$feature]
```

### AQI Data
```{r}
aqi_pollutants <- aqs_list_parameters(aqs_user, pc="AQI POLLUTANTS")
aqi_forecast <- aqs_list_parameters(aqs_user, pc="FORECAST")
```

```{r}
s2 <- aqs_dailyData(aqs_user=aqs_user,
                     endpoint="byCounty",
                     state=ny_fips_code,
                     county=nyCounty_fips_code,
                     bdate="20160101",
                     edate="20161231",
                     param=aqi_pollutants$code[1])
```

### Road Networks 
```{r}
# Read this shape file with the rgdal library. 
library(rgdal)

my_spdf <- readOGR( 
  dsn= paste0(getwd(), "/data/citymap_citymap_v1") , 
  layer="citymap_citymap_v1",
  verbose=FALSE)
```

```{r}
# Basic plot of this shape file:
par(mar=c(0,0,0,0))
plot(my_spdf, col="#f2f2f2", bg="black", lwd=0.25 )
```

```{r}
head(my_spdf)
```

```{r}
nyC_map <- my_spdf[my_spdf$boro_nm == BOROUGH]

is_(my_spdf$boro_nm == BOROUGH )

str(my_spdf)

names(my_spdf)

unique(my_spdf$ownership_)

my_spdf
```


```{r}
# Basic plot of this shape file:
par(mar=c(0,0,0,0))
plot(nyC_map, col="#f2f2f2", bg="black", lwd=0.25 )
```
