---
title: "Chicago: Master Dataset"
Author: Ally Racho
output: html_notebook
---

Notes:
- road network not included in feature table 
- had to average most variables since some datasets had multiple values per week
- rideshare / taxi date is given only as 1/1 so lots of NAs (removed for modeling)
- chicago traffic only 5 weeks in 2018 so again lots of NAs (removed for modeling)
- AQI missing a few weeks 

```{r}
#clear environment
rm(list = ls())
```

```{r}
#libraries need
library(readr)
library(sf)
library(tidyverse)
library(jsonlite)
library(lubridate)
```


## AQI
# data taken every 6 days
```{r}
#load data
ChicagoAQI <- read_csv("./Chicago/Data/ChicagoAQI.csv")

# split date into month / year 
ChicagoAQI <-
  ChicagoAQI %>%
  dplyr::mutate(year = lubridate::year(Data.date_local), 
                month = lubridate::month(Data.date_local), 
                day = lubridate::day(Data.date_local))

ChicagoAQI$week <- lubridate::week(ymd(ChicagoAQI$Data.date_local))

# subset out wanted variables and create new dataframe
vars <- c("Data.county", "Data.sample_measurement", "week", "month", "year")
ChicagoAQIcon <- na.omit(ChicagoAQI[vars])

# rename variables
ChicagoAQIcon <- 
  ChicagoAQIcon %>%
  rename(
    County = Data.county,
    AQI_Weekly_Measurement = Data.sample_measurement,
  )

ChicagoAQIcon <-
  ChicagoAQIcon %>%
  group_by(week, year) %>%
  mutate(AQI_Weekly_Measurement = mean(AQI_Weekly_Measurement)) %>%
  ungroup()

ChicagoAQIcon$City <- "Chicago"

ChicagoAQIcon <- distinct(ChicagoAQIcon)
ChicagoAQIcon <- distinct(ChicagoAQIcon, AQI_Weekly_Measurement, .keep_all = T)
write.csv(ChicagoAQIcon, file = "./Chicago/Data/Chicago_AQI.csv")
```

## WQI -- not used 
```{r}
#load data
ChicagoWQI <- read_csv("./Chicago/Data/ChicagoWQI.csv")

# reformat date before split
ChicagoWQI$ActivityStartDate <- strptime(as.character(ChicagoWQI$ActivityStartDate), "%m/%d/%Y")
format(ChicagoWQI$newdate, "%Y-%m-%d")

# split date into month / year 
ChicagoWQI <-
  ChicagoWQI %>%
  dplyr::mutate(year = lubridate::year(ActivityStartDate), 
                month = lubridate::month(ActivityStartDate), 
                day = lubridate::day(ActivityStartDate))

# subset out wanted variables and create new dataframe
vars <- c("ResultMeasureValue", "CharacteristicName", "month", "year", "ActivityLocation/LatitudeMeasure", "ActivityLocation/LongitudeMeasure", "ActivityStartTime/Time")
ChicagoWQIcon <- ChicagoWQI[vars]

# rename variables
ChicagoWQIcon <- 
  ChicagoWQIcon %>%
  rename(
    WQI_measure = ResultMeasureValue,
    WQI_month = month, 
    WQI_year = year,
    WQI_lat = 'ActivityLocation/LatitudeMeasure',
    WQI_long = 'ActivityLocation/LongitudeMeasure', 
    WQI_time = 'ActivityStartTime/Time'
  )

# add city name variable
ChicagoWQIcon$City <- "Chicago"
ChicagoWQIcon$County <- "Cook"
ChicagoWQIcon$WQI_year <- ChicagoWQIcon$WQI_year + 2000
```

## Weather

```{r}
#load data
chicagoWeather <- read_csv("./Chicago/Data/Chicago_weather.csv")


# split date into month / year 
chicagoWeather <-
  chicagoWeather %>%
  dplyr::mutate(year = lubridate::year(DATE), 
                month = lubridate::month(DATE), 
                day = lubridate::day(DATE))

chicagoWeather$week <- lubridate::week(ymd(chicagoWeather$DATE))


vars <- c("TMAX", "TMIN", "AWND", "PRCP", "week", "year", "month")
ChicagoWeathercon <- na.omit(chicagoWeather[vars])

# rename variables
ChicagoWeathercon <- 
  ChicagoWeathercon %>%
  rename(
    Weather_TMAX = TMAX, 
    Weather_TMIN = TMIN, 
    Weather_AWND = AWND, 
    Weather_PRCP = PRCP
  )


ChicagoWeathercon <- 
  ChicagoWeathercon %>%
  group_by(week, year) %>%
  mutate(Weather_TMAX = mean(Weather_TMAX), 
            Weather_TMIN = mean(Weather_TMIN),
            Weather_AWND = mean(Weather_AWND),
            Weather_PRCP = mean(Weather_PRCP))
ChicagoWeathercon <- distinct(ChicagoWeathercon)
ChicagoWeathercon <- distinct(ChicagoWeathercon, Weather_TMAX, .keep_all = T)
```

## Pollution
```{r}
#load data
chicagoPollution <- read_csv("./Chicago/Data/Chicago_pollution.csv")

# split date into month / year 
chicagoPollution <-
  chicagoPollution %>%
  dplyr::mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date))

chicagoPollution$week <- lubridate::week(ymd(chicagoPollution$date))

vars <- c("pm25", "o3", "year", "month", "week")
chicagoPollutioncon <- chicagoPollution[vars]

# rename variables
chicagoPollutioncon <- 
  chicagoPollutioncon %>%
  rename(
    Pollution_pm25 = pm25, 
    Pollution_o3 = o3
  )

#gather data to condense feature names and values
#chicagoPollutioncon <- pivot_longer(chicagoPollutioncon, cols = c("Pollution_pm25", "Pollution_o3"))
  

#only use data from 2018 - 2022
chicagoPollutioncon <- subset(chicagoPollutioncon, year >=2018)

ChicagoPollutioncon <- 
  chicagoPollutioncon %>%
  group_by(week, month, year) %>%
  summarise(Pollution_pm25 = mean(Pollution_pm25), 
            Pollution_o3 = mean(Pollution_o3))

```

## Traffic Congestion
```{r}
result <- fromJSON("./Chicago/Data/chitraffic.json")
chicagotraffic <- as.data.frame(result)

chicagotraffic <- separate(chicagotraffic, col = "time", into = c("date", "time"), sep = "T")

# split date into month / year 
chicagotraffic <-
  chicagotraffic %>%
  dplyr::mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date))

chicagotraffic$week <- lubridate::week(ymd(chicagotraffic$date))
chicagotraffic$bus_count <- as.numeric(chicagotraffic$bus_count)
chicagotraffic$speed <- as.numeric(chicagotraffic$speed)

vars <- c("week", "month", "year", "speed", "bus_count")
chicagotrafficcon <- na.omit(chicagotraffic[vars])


###
#gather data to condense feature names and values
#chicagotrafficcon <- pivot_longer(chicagotrafficcon, cols = c("traffic_speed", "traffic_bus_count"))

chicagotrafficcon <- 
  chicagotrafficcon %>%
  rename(
    traffic_speed = speed,
    traffic_bus_count = bus_count,
  )

#chicagotrafficcon$feature_lat <- as.numeric(chicagotrafficcon$feature_lat)
#chicagotrafficcon$feature_long <- as.numeric(chicagotrafficcon$feature_long)
#chicagotrafficcon$feature_val <- as.numeric(chicagotrafficcon$feature_val)

chicagotrafficcon <- 
  chicagotrafficcon %>%
  group_by(week, month, year) %>%
  summarise(traffic_speed = mean(traffic_speed), 
            traffic_bus_count = mean(traffic_bus_count))
```

## Taxi Traffic
```{r}
result <- fromJSON("./Chicago/Data/chi_taxi2018.json")
chicagotaxi2018 <- as.data.frame(result)

chicagotaxi2018 <- subset(chicagotaxi2018, select = -c(pickup_centroid_location, dropoff_centroid_location))
result <- fromJSON("./Chicago/Data/chi_taxi2019.json")
chicagotaxi2019 <- as.data.frame(result)
chicagotaxi2019 <- subset(chicagotaxi2019, select = -c(pickup_centroid_location, dropoff_centroid_location))
result <- fromJSON("./Chicago/Data/chi_taxi2020.json")
chicagotaxi2020 <- as.data.frame(result)
chicagotaxi2020 <- subset(chicagotaxi2020, select = -c(pickup_centroid_location, dropoff_centroid_location))
result <- fromJSON("./Chicago/Data/chi_taxi2021.json")
chicagotaxi2021 <- as.data.frame(result)
chicagotaxi2021 <- subset(chicagotaxi2021, select = -c(pickup_centroid_location, dropoff_centroid_location))
#combine all years to one dataset
df_list <- list(chicagotaxi2018, chicagotaxi2019, chicagotaxi2020, chicagotaxi2021)

###### fix -- duplicate 'row.names' are not allowed
ChicagoTaxi <- Reduce(function(x,y)merge(x,y,all = TRUE), df_list, accumulate = FALSE)

# split date into month / year 
ChicagoTaxi <-
  ChicagoTaxi %>%
  dplyr::mutate(year = lubridate::year(trip_start_timestamp), 
                month = lubridate::month(trip_start_timestamp), 
                day = lubridate::day(trip_start_timestamp))
ChicagoTaxi$trip_miles <- as.numeric(ChicagoTaxi$trip_miles)

ChicagoTaxi$trip_total <- as.numeric(ChicagoTaxi$trip_total)

ChicagoTaxi$week <- 1


#select vars
vars <- c("year", "week", "month", "trip_miles", "trip_total")

#subset
chicagoTaxicon <-  ChicagoTaxi[vars]
# rename variables
chicagoTaxicon <- 
  chicagoTaxicon %>%
  rename(
    traffic_taxi_trip_miles = trip_miles, 
    traffic_taxi_trip_totals = trip_total
  )

chicagoTaxicon <- 
  chicagoTaxicon %>%
  group_by(year, month, week) %>%
  summarise(traffic_taxi_trip_miles = mean(traffic_taxi_trip_miles), 
            traffic_taxi_trip_totals = mean(traffic_taxi_trip_totals))
```

## Rideshare
```{r}
result <- fromJSON("./Chicago/Data/ridesharechi.json")
chicagoRideshare <- as.data.frame(result)


#select vars
vars <- c("trip_start_timestamp", "trip_miles", "dropoff_centroid_longitude", 
          "dropoff_centroid_latitude")


# split date into month / year 
chicagoRideshare <-
  chicagoRideshare %>%
  dplyr::mutate(year = lubridate::year(trip_start_timestamp), 
                month = lubridate::month(trip_start_timestamp), 
                day = lubridate::day(trip_start_timestamp))
chicagoRideshare$week <- 1

#select vars
vars <- c("month", "week", "trip_miles", "year")

#subset
chicagoRidesharecon <- chicagoRideshare[vars]
chicagoRidesharecon$trip_miles <- as.numeric(chicagoRidesharecon$trip_miles)
chicagoRidesharecon$rides_total <- chicagoRidesharecon$trip_miles - sum(chicagoRidesharecon$trip_miles)


# rename variables
chicagoRidesharecon <- 
  chicagoRidesharecon %>%
  rename(
    traffic_rideshare_miles = trip_miles)



chicagoRidesharecon <- subset(chicagoRidesharecon, year >=2018)

chicagoRidesharecon <- 
  chicagoRidesharecon %>%
  group_by(week,month,year) %>%
  summarise(traffic_rideshare_miles = mean(traffic_rideshare_miles))

```

## Bus Routes Daily Totals (no lat  / long), 2018 - 2020
```{r}
result <- read.csv("./Chicago/Data/Chicago_bus.csv")
chicagoBus <- as.data.frame(result)

# reformat date before split
chicagoBus$Month_Beginning <- strptime(as.character(chicagoBus$Month_Beginning), "%m/%d/%Y")
format(chicagoBus$Month_Beginning, "%Y-%m-%d")

# split date into month / year 
chicagoBus <-
  chicagoBus %>%
  dplyr::mutate(year = lubridate::year(Month_Beginning), 
                month = lubridate::month(Month_Beginning), 
                day = lubridate::day(Month_Beginning))

chicagoBus$week <- lubridate::week(ymd(chicagoBus$Month_Beginning))

#select vars
vars <- c("month","week", "Avg_Weekday_Rides", "year", "Avg_Saturday_Rides", "Avg_Sunday.Holiday_Rides", "MonthTotal")

#subset
chicagoBuscon <- chicagoBus[vars]
chicagoBuscon$avg_weekday_rides <- as.numeric(chicagoBuscon$Avg_Weekday_Rides)
chicagoBuscon$avg_saturday_rides <- as.numeric(chicagoBuscon$Avg_Saturday_Rides)
chicagoBuscon$avg_sunday_holiday_rides <- as.numeric(chicagoBuscon$Avg_Sunday.Holiday_Rides)



chicagoBuscon <- subset(chicagoBuscon, year >=2018)
vars <- c("month", "week", "year", "avg_weekday_rides", "avg_saturday_rides", "avg_sunday_holiday_rides")
chicagoBuscon <- chicagoBuscon[vars]

chicagoBuscon <- 
  chicagoBuscon %>%
  group_by(week, month, year) %>%
  summarise(avg_weekday_rides = mean(avg_weekday_rides), 
            avg_saturday_rides = mean(avg_saturday_rides), 
            avg_sunday_holiday_rides = mean(avg_sunday_holiday_rides))
```

## Road Network -- not included in feature set
```{r}
library(rgdal)
Chicago_spdf <- readOGR( 
  dsn= paste0(getwd(),"./Chicago/Data/cityboundary/cityboundary.shp"))

par(mar=c(0,0,0,0))
plot(Chicago_spdf, col="#f2f2f2", bg="black", lwd=0.25, border = 0)
```


## Combine datasets to create master
```{r}
# combine row feature datasets
chi_v1 <- merge(ChicagoAQIcon, ChicagoWeathercon, by = c("week", "month", "year"), all.y= T)
chi_v12 <- merge(chi_v1, chicagoBuscon, by = c("week", "month", "year"), all.x= T)
chi_v2 <- merge(chi_v12, ChicagoPollutioncon, by = c("week", "month", "year"), all.x = T)
chi_v3 <- merge(chi_v2, chicagoTaxicon, by = c("week", "month", "year"), all.x = T)
chi_v4 <- merge(chi_v3, chicagoRidesharecon, by = c("week", "month", "year"), all.x = T)

Chicago_feature <- merge(chi_v4, chicagotrafficcon, by = c("week", "month", "year"), all.x = T)
write.csv(Chicago_feature, file = "./Chicago/Data/ChicagoFeatures_V2.csv")
```
