---
title: "Philadelphia: Master Dataset"
Author: Ally Racho
output: html_notebook
---

## Notes:
- AQI is 3 day span for county of Philadelphia, Philadelphia. -- take avg over 6 days
- road network not included in feature set 
- still dealing with lots of NAs 



# clear environment
```{r}
#clear environment
rm(list = ls())
```

#libraries need
```{r}
library(jsonlite)
library(readr)
library(lubridate)
library(dplyr)
library(tidyverse)
library(sf) # shp file 
```

## AQI
# data taken every 3 days, took avg of week
```{r}
#load data
PhillyAQI <- read_csv("./Data/PhillyAQI.csv")

# split date into month / year 
PhillyAQI <-
  PhillyAQI %>%
  dplyr::mutate(year = lubridate::year(Data.date_local), 
                month = lubridate::month(Data.date_local), 
                day = lubridate::day(Data.date_local))

PhillyAQI$week <- lubridate::week(ymd(PhillyAQI$Data.date_local))

# subset out wanted variables and create new dataframe
vars <- c("Data.county", "Data.sample_measurement","week", "month", "year")
PhillyAQIcon <- na.omit(PhillyAQI[vars])

# rename variables
PhillyAQIcon <- 
  PhillyAQIcon %>%
  rename(
    County = Data.county,
    AQI_values = Data.sample_measurement,
  )

# add city name variable
PhillyAQIcon$City <- "Philadelphia"

vars <- c("County", "City", "AQI_values", "week", "month", "year")
PhillyAQIcon <- PhillyAQIcon[vars]

PhillyAQIcon <-
  PhillyAQIcon %>%
  group_by(week, year) %>%
  mutate(AQI_Weekly_Measurement = mean(AQI_values)) 

PhillyAQIcon <- distinct(PhillyAQIcon)
PhillyAQIcon <- distinct(PhillyAQIcon, AQI_Weekly_Measurement, .keep_all = T)
vars <- c("County", "City", "AQI_Weekly_Measurement", "week", "month", "year")
PhillyAQIcon <- PhillyAQIcon[vars]

write.csv(PhillyAQIcon, "./Data/Philly_AQI.csv")
```


## WQI 
# need to choose which features needed
```{r}
# load data
PhillyWQI <- read_csv("./Data/PhillyWQI.csv")

# reformat date before split
PhillyWQI$ActivityStartDate <- strptime(as.character(PhillyWQI$ActivityStartDate), "%m/%d/%Y")
format(PhillyWQI$newdate, "%Y-%m-%d")

# split date into month / year 
PhillyWQI <-
  PhillyWQI %>%
  dplyr::mutate(year = lubridate::year(ActivityStartDate), 
                month = lubridate::month(ActivityStartDate), 
                day = lubridate::day(ActivityStartDate))

# subset out wanted variables and create new dataframe
vars <- c("ResultMeasureValue", "CharacteristicName", "month", "year", "ActivityLocation/LatitudeMeasure", "ActivityLocation/LongitudeMeasure", "ActivityStartTime/Time")
PhillyWQIcon <- PhillyWQI[vars]

# rename variables
PhillyWQIcon <- 
  PhillyWQIcon %>%
  rename(
    WQI_measure = ResultMeasureValue,
    WQI_month = month, 
    WQI_year = year,
    WQI_lat = 'ActivityLocation/LatitudeMeasure',
    WQI_long = 'ActivityLocation/LongitudeMeasure', 
    WQI_time = 'ActivityStartTime/Time'
  )

# add city name variable
PhillyWQIcon$City <- "Philadelphia"
PhillyWQIcon$County <- "Philadelphia"
PhillyWQIcon$WQI_year <- PhillyWQIcon$WQI_year + 2000
```


# Weather
```{r}
#load data
phillyWeather <- read_csv("./Data/Philly_weather.csv")

# split date into month / year 
phillyWeather <-
  phillyWeather %>%
  dplyr::mutate(year = lubridate::year(DATE), 
                month = lubridate::month(DATE), 
                day = lubridate::day(DATE))

phillyWeather$week <- lubridate::week(ymd(phillyWeather$DATE))

vars <- c( "TMAX", "TMIN", "AWND", "PRCP","week", "year", "month")
PhillyWeathercon <- na.omit(phillyWeather[vars])

# rename variables
PhillyWeathercon <- 
  PhillyWeathercon %>%
  rename(
    Weather_TMAX = TMAX, 
    Weather_TMIN = TMIN, 
    Weather_AWND = AWND, 
    Weather_PRCP = PRCP
  )

PhillyWeathercon <- 
  PhillyWeathercon %>%
  group_by(week, year) %>%
  mutate(Weather_TMAX = mean(Weather_TMAX), 
            Weather_TMIN = mean(Weather_TMIN),
            Weather_AWND = mean(Weather_AWND),
            Weather_PRCP = mean(Weather_PRCP))
PhillyWeathercon <- distinct(PhillyWeathercon)
PhillyWeathercon <- distinct(PhillyWeathercon, Weather_TMAX, .keep_all = T)
```

## Pollution
```{r}
#load data
phillyPollution <- read_csv("./Data/Philly_pollution.csv")

# split date into month / year 
phillyPollution <-
  phillyPollution %>%
  dplyr::mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date))

phillyPollution$week <- lubridate::week(ymd(phillyPollution$date))


vars <- c("pm25", "o3", "no2", "so2", "co", "year", "month", "week")
phillyPollutioncon <- phillyPollution[vars]

# rename variables
phillyPollutioncon <- 
  phillyPollutioncon %>%
  rename(
    Pollution_pm25 = pm25, 
    Pollution_o3 = o3, 
    Pollution_no2 = no2, 
    Pollution_so2 = so2, 
    Pollution_co = co
  )

# only use data from 2018 - 2021
phillyPollutioncon <- subset(phillyPollutioncon, year>=2018)
phillyPollutioncon <- subset(phillyPollutioncon, year<2022)

phillyPollutioncon <- 
  phillyPollutioncon %>%
  group_by(week,month, year) %>%
  summarise(Pollution_pm25 = mean(Pollution_pm25), 
            Pollution_o3 = mean(Pollution_o3),
            Pollution_no2 = mean(Pollution_no2),
            Pollution_so2 = mean(Pollution_so2), 
            Pollution_co= mean(Pollution_co))

phillyPollutioncon <- phillyPollutioncon[-c(21:24, 41:44, 61:63, 84:87, 104:107, 124:126, 147:150,148, 158, 167:170,203:205), ]

```


## Vehicle Counts
```{r}
Phillytraffic <- read_csv("./Data/Greater_Philadelphia_Vehicle_Volume_Counts.csv")

# split date into month / year 
Phillytraffic <-
  Phillytraffic %>%
  dplyr::mutate(year = lubridate::year(setdate), 
                month = lubridate::month(setdate), 
                day = lubridate::day(setdate))

Phillytraffic$week <- week(Phillytraffic$setdate)

vars <- c("year","week", "month", "co_name", "aadt")
PhillyTrafficcon <- Phillytraffic[vars]

# rename variables
PhillyTrafficcon <- 
  PhillyTrafficcon %>%
  rename(
    traffic_vehicle_counts = aadt,
  )

# only use Philadelphia County data
PhillyTrafficcon <- subset(PhillyTrafficcon, (co_name =="Philadelphia"))
PhillyTrafficcon <- subset(PhillyTrafficcon, (year >= 2018))

PhillyTrafficcon <- subset(PhillyTrafficcon, select = -c(co_name))

PhillyTrafficcon <- 
  PhillyTrafficcon %>%
  group_by(week, month, year) %>%
  summarise(traffic_vehicle_counts = mean(traffic_vehicle_counts))
```

## Bicycle Counts
```{r}
PhillyBike <- read_csv("./Data/Greater_Philadelphia_Bicycle_Counts.csv")

# split date into month / year 
PhillyBike <-
  PhillyBike %>%
  dplyr::mutate(year = lubridate::year(setdate), 
                month = lubridate::month(setdate), 
                day = lubridate::day(setdate))

PhillyBike$week <- week(PhillyBike$setdate)

vars <- c("year","week", "month", "co_name", "aadb")
PhillyBikecon <- PhillyBike[vars]

# rename variables
PhillyBikecon <- 
  PhillyBikecon %>%
  rename(
    traffic_bike_counts = aadb,
  )

# only use Philadelphia County data
PhillyBikecon <- subset(PhillyBikecon, (co_name =="Philadelphia"))
PhillyBikecon <- subset(PhillyBikecon, (year >= 2018))

PhillyBikecon <- subset(PhillyBikecon, select = -c(co_name))

PhillyBikecon <- 
  PhillyBikecon %>%
  group_by(week, month, year) %>%
  summarise(traffic_bike_counts = mean(traffic_bike_counts))
                      
```


## Pedestrian Count
```{r}
PhillyPed <- read_csv("./Data/Greater_Philadelphia_Pedestrian_Counts.csv")

# split date into month / year 
PhillyPed <-
  PhillyPed %>%
  dplyr::mutate(year = lubridate::year(setdate), 
                month = lubridate::month(setdate), 
                day = lubridate::day(setdate))

PhillyPed$week <- week(PhillyPed$setdate)

vars <- c("year", "week", "month", "co_name", "aadp")
PhillyPedcon <- PhillyPed[vars]

# rename variables
PhillyPedcon <- 
  PhillyPedcon %>%
  rename(
    traffic_ped_counts = aadp,
  )

# only use Philadelphia County data
PhillyPedcon <- subset(PhillyPedcon, (co_name =="Philadelphia"))
PhillyPedcon <- subset(PhillyPedcon, (year >= 2018))

PhillyPedcon <- subset(PhillyPedcon, select = -c(co_name))

PhillyPedcon <- 
  PhillyPedcon %>%
  group_by(week, month, year) %>%
  summarise(traffic_ped_counts = mean(traffic_ped_counts))
```

## Road Network
```{r}
#library(rgdal)
#Philly_spdf <- readOGR( 
  #dsn= paste0(getwd(),"./Philly/Data/Philly_map.shp"), 
  #layer = "lines",
  #verbose = FALSE
#)

#par(mar=c(0,0,0,0))
#plot(Philly_spdf, col="#f2f2f2", bg="black", lwd=0.25, border=0 )
```


## Combine datasets to create master
```{r}
# combine row feature datasets
PHL_v1 <- merge(PhillyAQIcon, PhillyWeathercon, by = c("week", "month", "year"), all.y = T)
PHL_v12 <- merge(PHL_v1, PhillyBikecon, by = c("week", "month", "year"), all.x= T)
PHL_v2 <- merge(PHL_v12, phillyPollutioncon, by = c("week", "month", "year"), all.x = T)
PHL_v3 <- merge(PHL_v2, PhillyPedcon, by = c("week", "month", "year"), all.x = T)
Philly_feature <- merge(PHL_v3, PhillyTrafficcon, by = c("week", "month", "year"), all.x = T)
write.csv(Philly_feature, file = "./Data/PhillyFeatures_V2.csv")
```
